---
title: "Adding Tidiers to Broom"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Adding new tidiers to broom

Thank you for your interest in adding new tidiers to broom! This document is a *work in progress* describing the conventions that broom follows.

Some general guidelines:

- Try to reach a minimum 90% test coverage for new tidiers
- `tidy`, `glance` and `augment` methods *must* return tibbles
- Use the [tidyverse style conventions](http://style.tidyverse.org/)
- Feel free to add yourself as a contributor to `DESCRIPTION`

<!-- TODOS: -->

<!-- - Helper functions that you should be aware of and use -->
<!-- - Preferred packages to use: i.e. new tidyverse not old -->

<!-- Comments about tidiers living in external packages -->

# `tidy`

Each row in a `tidy` output provides information about the components of a model, such as one term in a regression, one test, or one cluster/class. This meaning varies across models but is usually self-evident.

## Column names to use in output

* `term`: the term in a regression or model that is being estimated.
* `p.value`: the p-value associated with a test statistic
* `statistic`: a test statistic, usually the one used to compute the p-value. Combining these across many sub-groups is a reliable way to perform (e.g.) bootstrap hypothesis testing
* `estimate`: estimate of an effect size, slope, or other value
* `std.error`: standard error
* `conf.low`: the low end of a confidence interval on the `estimate`
* `conf.high`: the high end of a confidence interval on the `estimate`
* `df`: degrees of freedom


# `glance`

The `glance` method accepts a model object and returns a tibble with a single row of model level summaries.

Do not include arguments *given* to the modeling function. For example, a `glm` glance output does not need to contain a field for `family`, since that is decided by the user calling `glm` rather than the modeling function itself.

## Column names to use in output:

* `r.squared` the fraction of variance explained by the model
* `adj.r.squared` R^2 adjusted based on the degrees of freedom
* `sigma` the square root of the estimated variance of the residuals


# `augment`

The `augment` method accepts a model object and optionally a data frame and returns a tibble of data with additional observation level information. `augment` should be used for information related to individual data points.

* `augment(model, data)` adds columns to the original data.
    * If the `data` argument is missing, `augment` attempts to reconstruct the data from the model (note that this may not always be possible, and usually won't contain columns not used in the model).

`augment(object)` should return a tibble of summaries for the original observations.

`augment(object, newdata = newdata)` should return a tibble of observation level summaries for a new data frame.

Each row in `augment(object)` should correspond to a single row of the original data, or a single row of `newdata` if `newdata` was passed as an argument.

If the original data contained rownames, `augment` should return them in a column called `.rownames`.

Newly added column names begin with `.` to avoid overwriting columns in the original data.

## Arguments that may be relevant:

- x: object to augment
- type.residuals
- type.predict
- data
- newdata
- weights


## Column names to use in output

* `.fitted`: the predicted values, on the same scale as the data.
* `.resid`: residuals: the actual response values minus the fitted values
* `.cluster`: cluster assignments


# Testing new tidiers

## Basic checks

All `tidy` methods should `check_tidy`, all `glance` methods should pass `check_glance` and all `augment` methods should pass `check_augment`. That is, your test should include:

```{r, eval = FALSE}
t <- tidy(object)
g <- glance(object)
a <- augment(object)

check_tidy(t)
check_glance(g)
check_augment(a)
```

## Package dependencies

In order to test your tidiers, you may need to add `your_package` to the **Suggested** section of broom's DESCRIPTION.

Here's a outline that you can use for your test file

```{r}
# TODO. requireNamespace(package, quietly = TRUE)?
```

At the moment, pull requests must pass the AppVeyor build to be added to `broom`. It is okay if the Travis CI build times out.

## Catching edge cases

Many modelling function allow users to specify a variety of options. For example, `glm` allows various response types via the `family` argument. Good tests for `glm` tidying methods would test tidiers on `glm` objects fit for all acceptable values of `family`.

# Before submitting your pull request

- Make sure that you have run and passed:
    - `devtools::spell_check()`
    - `goodpractice::gp()`
- Run `pkgdown::build_site()` so that your documentation gets added to the broom
- R CMD check

An effective pull request should contain a big picture summary of what you've done and a brief example usage in the pull request text.
