---
title: "Exporting tidying methods from a package"
author: "Alex Hayes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exporting tidying methods from a package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

If you are a package developer, the best place to include `tidy()`, `glance()` or `augment()` methods is in your own package. To do this, you'll need to:

1. Re-export the `tidy()`, `glance()` and `augment()` generics from the `modelgenerics` package.
2. Implement appropriate tidying methods.
3. Tests the tidying methods

The [dustpan](https://github.com/alexpghayes/dustpan) package demonstrates use of `modelgenerics` and `modeltests`, and you find it helpful to browse.

## Re-exporting `tidy()`, `glance()` and `augment()` generics

First you'll need to add the [modelgenerics](https://github.com/tidymodels/modelgenerics) package to `Imports`. We recommend using the [usethis](https://github.com/r-lib/usethis) package for this:

```{r}
usethis::use_package("modelgenerics", "Imports")
```

Next you'll need to re-export the appropriate tidying methods. If you plan to implement a `glance()` method, for example, you can re-export the glance generic by adding the following somewhere in your package

```{r}
#' @importFrom modelgenerics glance
#' @export
modelgenerics::glance
```

Run `devtools::document()` for these changes to take effect. 

Note: please do not define `tidy()`, `glance()` or `augment()` generics in your package. This will result in namespace conflicts whenever your package is used along other packages that also export tidying methods.

## Implement appropriate tidying methods

This part is mostly up to you, although there's lots of advice in `vignette("adding-tidiers")`.

## Testing

Now that you have a tidying method, you'll want to test it. In addition to testing the specific functionality of your tidying methods, you should also pass the standardized tests exported from [modeltests](https://github.com/alexpghayes/modeltests).

To import these tests you'll need to add `modeltests` to `Suggests`, which you can accomplish with:

```{r}
# once modeltests makes its way onto CRAN:
# usethis::use_package("modeltests", "Suggests")

# until then:
remotes::install_github("alexpghayes/modeltests")
usethis::use_dev_package("modeltests")
```

If you're using [testthat](https://github.com/r-lib/testthat), you'll want to include the following in your tests:

```{r}
skip_if_not_installed("modeltests")  # can't rely on suggested packages in tests
library(modeltests)

test_that("myobject tidier arguments", {
  check_arguments(tidy.my_object)
  check_arguments(glance.my_object)
  check_arguments(augment.myobject)
})

test_that("tidy.my_object", {
  my_object <-  my_object(y ~ x, data = my_data, ...)
  
  td <- tidy(my_object)
  tdci <- tidy(my_object, conf.int = TRUE)
  # ...
  
  check_tidy_output(td)
  check_tidy_output(tdci)
  # ...
})

test_that("glance.my_object", {
  my_object <-  my_object(y ~ x, data = my_data, ...)
  my_object2 <-  my_object(y ~ x, data = my_data2, ...)
  
  gl <- glance(my_object)
  gl2 <- glance(my_object2)
  
  check_glance_outputs(gl, gl2)
})

test_that("augment.my_object", {
  my_object <-  my_object(y ~ x, data = my_data, ...)
  
  check_augment_function(
    aug = augment.my_object,
    model = my_object,
    data  = my_data,
    newdata = my_new_data  # or just my_data
  )
})

```

If you're run `devtools::test()` and `devtools::check()` and didn't fail any tests or set off R CMD Check ERRORS, WARNINGS or NOTES, you're done!

### If you're tidiers don't pass the tests

If you are introducing new arguments or column names in output, you'll need to make a pull request to `modeltests` to add the new argument/column names to the glossary of allowed names.

If you'd like to change the test specifications, or if you find a bug in the tests, open an issue in `modeltests`.

As a last resort, you can set `strict = FALSE` to run only a bare minimum of tests. This is not recommended.
